---
title: "EDA of Data Snippet"
author: "Connor Gibbs"
header-includes:
   - \usepackage{amsmath}
   - \usepackage{bm}
   - \usepackage{mathtools}
   - \usepackage{braket}
date: "10/02/2020"
output: 
   bookdown::pdf_document2:
    fig_caption: yes
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.width=6, fig.height=3, fig.align="center") 
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
# dependencies
library(tidyverse)
library(janitor)
library(jsonlite)
library(dtplyr)
library(igraph)
library(ggnetwork)
```

# Integers and Node Labels

The integer data consists of three columns: subject, predicate, and object. The three columns allow for the specification of a property graph where subject and object are nodes of the graph and the predicate specifies the value of the edge between the nodes.

The node label data consists of four columns: node_id, label, description/definition, and synonym. The node_id seems to link the integer data to the node label data. The node_id seems to be the unique identifier for the subject and object from the integer data, providing further details on these nodes. 

```{r}
# integers
int_map <- './hpo-gene/src/data/PheKnowLator_Instance_RelationsOnly_NotClosed_NoOWL_Triples_Integers.txt' %>%
  read_tsv() %>%
  clean_names()
head(int_map)

# node labels
node_labels <- './hpo-gene/src/data/PheKnowLator_Instance_RelationsOnly_NotClosed_NoOWL_NodeLabels.txt' %>%
  read_tsv() %>%
  clean_names()
head(node_labels)

# identifiers
ids <- './hpo-gene/src/data/PheKnowLator_Instance_RelationsOnly_NotClosed_NoOWL_Triples_Identifiers.txt' %>%
  read_tsv() %>%
  clean_names()
head(ids)

json_map <- './hpo-gene/src/data/PheKnowLator_Instance_RelationsOnly_NotClosed_NoOWL_Triples_Integer_Identifier_Map.json' %>%
  read_json()
head(json_map[[1]])

json_map <- data.frame(origin = names(json_map), values = unlist(unname(json_map)))

int_map <- int_map %>%
  pivot_longer(., cols = everything(), names_to = 'type', values_to = 'tnodeID')
  
my_map <- ids %>%
  pivot_longer(., cols = everything(), names_to = 'type', values_to = 'tnodeLabel') %>%
  add_column(., tnodeID = int_map$tnodeID) %>%
  mutate(., nodeLabel = basename(tnodeLabel)) %>%
  select(., nodeLabel, tnodeLabel, tnodeID) %>%
  distinct()
write_csv(my_map, 'hpo-gene/src/data/full_set/map2full.csv')
```

Let's combine these two data sets into one, removing the old data sets to preserve memory.

```{r, include=FALSE}
prop_graph <- int_map %>%
  left_join(., node_labels, by = c('subject' = 'node_id')) %>%
  rename_at(., vars(label, description_definition, synonym), ~str_c('subject.', .x)) %>%
  left_join(., node_labels, by = c('object' = 'node_id')) %>%
  rename_at(., vars(label, description_definition, synonym), ~str_c('object.', .x)) %>%
  select(., starts_with('subject'), 'predicate', starts_with('object'))

rm(int_map, node_labels)
```

Now, let's see how many edges there are for nodes with the top ten most edges.

```{r}
subject_counts <- prop_graph %>%
  lazy_dt() %>%
  group_by(., subject) %>%
  summarize(., n = n()) %>%
  ungroup() %>%
  arrange(., desc(n)) %>%
  as_tibble()

subject_counts %>%
  slice(., 1:10) %>%
  left_join(., distinct(prop_graph, subject, subject.label)) %>%
  select(., subject, subject.label, n)
```

To get sense of the graphical structure, let's consider twenty subjects with relatively few edges:

```{r, cache=TRUE}
topn_graph <- prop_graph %>%
  filter(., subject %in% subject_counts$subject[100000:100020]) %>%
  arrange(., object, predicate)

net <- topn_graph %>%
  select(., from = subject, to = object) %>%
  graph.data.frame(., directed = FALSE)

ggplot(net, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "grey50") +
  geom_nodes() +
  theme_blank()
```

Now, let's consider the solution to the diffusion kernel:

```{r}
library(Matrix)
A <- as_adjacency_matrix(net)
D <- diag(colSums(A))
W <- D - A
expm(-0.05*W)[1:5,1:5]
```











